### 2019-10-15 tesis_hiram/Chapter1/chapter1.tex
algo

 Quarks combine to form composite particles called hadrons. Quarks cannot be found in isolation, they always exist in form of hadrons that can be mesons or baryons. \\
Mesons , are bosons made up of a quark-antiquark pair, such as pion ($\pi$) and kaon ($K$).Mesons were theorized by Yukawa in 1937 and discovered in 1947 by Powell. \cite{griff}
Baryons are fermions made of 3 quarks. Examples of baryons are protons and neutrons. Both particles can interact via strong force.
\\

@@

\section{Top quark}
The top quark  is the most massive particle  with a charge of 2/3 with spin of 1/2. First proposed by  M. Kobayashi and T. Maskawa in 1973 to explain observed CP violations in Kaon decay\cite{griff}. On 1995, in the Tevatron at Fermilab in Illinois, the top quark  was discovered via proton antiproton collisions with center of mass energy between 1.8 to 1.96 TeV in the CDF and D0 experiments.The top quark production process was via ggF, which generated a pair of $t\bar{t}$\cite{top} .
Since then,the properties of the top quark have been studied at the Tevatron Collider at Fermilab, and by the ATLAS and CMS experiments at the Large Hadron Collider (LHC) at CERN.\\
However, the top quark often receives special attention in new physics models because its mass requires near unity coupling to the Higgs boson \cite{top}.
As shown in table \ref{SM table}, the top quark lifetime is so small that they decay before they can hadronise, that is to say, the process to create hadrons from quarks and gluons.\\

@@

The production of top quark can be generated in differents ways. In tevatron and LHC, the production of $t\bar{t}$ comes from ggF (gluon gluon fusion) via strong force. Even it is posible to create a single top quark in the LHC via electroweak interactions (with a W boson) \cite{th1}.  
The top quark, due to its big mass, always decays into a W boson and a b quark. But since the W boson have various decays, the top quark can generate many posible particles, acording to W boson decays. Some posible decays of top quark is the decay to a lepton such as muon, electron or tau and their neutrinos . %The antitop, antiparticle of top quark, can decay to W$^{-1} \bar{b}$ and its decay is the same as top, but with inverse signal.

@@

$W^i$ (i=1,2,3) are three  SU(2) gauge bosons and B is a U(1) gauge boson.\footnote{$\epsilon^{ijk}$ represents the Levi-Civita tensor, where $\epsilon^{ijk}= \begin{dcases}
	+1 \quad  \text{for even permutation} \\
	-1 \quad \text{for odd permutation}\\
	0 \quad \text{i=j, or j=k, or k=i } 
	\end{dcases}$.} In the end, those gauge bosons, after introduce the Higgs mechanism, become the $W^\pm$ , $Z$ and $\gamma$ vector boson \cite{ew1,ew2}. %Z bosons  and photons will be formed by a combination of W and B fields.\\
\\

@@

In the recent years, people have studied the top anti top Higgs  ($t\bar{t}H$) process, in order to detect a Higgs boson, via multilepton signals \cite{th1}.
In this process, the creation of a Higgs boson comes from a intection via $t\bar{t}$. In the figure \ref{psu}, the $t\bar{t}H$ starts with 2 gluons which decay to two $t\bar{t}$ pairs and after that , the interaction $t\bar{t}$  generates a Higgs. The last process, subject of study in this thesis, is the top Higgs ($tH$) process. This process is a rare expected process with a very small production rate\cite{pd}.  In the figure \ref{thw}, is another way to produce $tH$, in association with a $W$.

@@

The number of particles decreases over time exponentially and this can be measured. With this, it is possible determinate the mean lifetime\cite{griff}
\begin{align}
\tau=\frac{1}{\Gamma}
\end{align}
In case of  the decay rate to a specific process, it is necessary get the branching fraction of the process. 
The branching ratio for a decay process is the ratio of the number of particles which decay via a specific decay mode with respect to the total number of particles which decay via all decay modes.
\begin{align}
BR_i =\frac{\Gamma_i}{\sum_{i}\Gamma_i}
\end{align}
Where $\Gamma=\sum_i\Gamma_i$ is the total decay width (sum of all partial widths) of the particle and is related to lifetime of the particle: $\Gamma=1/\tau$.
Since the dimension of $\Gamma$ is the inverse of time, in our system of natural units, it is measured in inverse seconds,it has the same dimension as mass (or energy)\footnote{Cleaves H.J. (2011) Branching Ratio. In: Gargaud M. et al. (eds) Encyclopedia of Astrobiology. Springer, Berlin, Heidelberg} . The lifetime of the Higgs boson is predicted to be 1.56 $\times$ $10^{-22}$ seconds and corresponds to $\Gamma_H$ is about 4 MeV , but has not yet been measured due to the detector resolution \cite{cms-manual}.
\\
\begin{table}[!htbp] 
\caption{SM Higgs boson branching ratios for  $M_H$ =125 GeV \protect \cite{pd}}
\centering
\begin{tabular}{|c|c|}
\hline
Higgs decay & Branching ratio (BR)\\
\hline
$H \rightarrow$ b$\bar{b}$ &$58.4\%$ \\
\hline
 $H \rightarrow$ $W^+W^-$ &$21.4\%$ \\
\hline
$H \rightarrow$ $\tau^+ \tau^-$ & $6.27\%$\\
\hline
$H \rightarrow$ ZZ &$2.62\%$\\
\hline
$H \rightarrow$ $\gamma\gamma$ &$0.227\%$\\
\hline
$H \rightarrow$ Z$\gamma$ &$0.153\%$\\
\hline
$H \rightarrow$ $\mu^+\mu^-$ &$0.0218\%$\\
\hline
\end{tabular}
\label{higgs1}
\end{table}
The decay  $H \rightarrow$ $b\bar{b}$ have a branching ratio a bit more than 50$\%$, therefore is the most likely decay Higgs boson can decay. Due to this,  $H\rightarrow b\bar{b}$ is a very studied process and most likely to detect a Higgs boson in the data after reconstruction. The decay  $H \rightarrow$ $\mu^+\mu^-$ has a very low branching ratio. This decay is very rare but not imposible to detect in the future.  In this work the majority of the Higgs decays detected are from $WW$, $ZZ$, $\tau \tau$\\

>>>>

 Quarks combine to form composite particles called hadrons. Quarks cannot be found in isolation, they always exist in form of hadrons that can be mesons or baryons. \\
Mesons , are bosons made up of a quark-antiquark pair, such as pion ($\pi$) and kaon ($K$).Mesons were theorized by Yukawa in 1937 and discovered in 1947 by Powell\cite{griff}.
Baryons are fermions made of 3 quarks. Examples of baryons are protons and neutrons. Both particles can interact via strong force.
\\

@@

\section{Top quark}
The top quark  is the most massive particle  with a charge of 2/3 with spin of 1/2. First proposed by  M. Kobayashi and T. Maskawa in 1973 to explain observed CP violations in Kaon decay\cite{griff}. On 1995, in the Tevatron at Fermilab in Illinois, the top quark  was discovered via proton antiproton collisions with center of mass energy between 1.8 to 1.96 TeV in the CDF and D0 experiments.The top quark production process was via ggF, which generated a pair of $t\bar{t}$\cite{top} .
Since then,the properties of the top quark have been studied at the Tevatron Collider at Fermilab, and by the ATLAS and CMS experiments at the Large Hadron Collider (LHC) at CERN.\\
However, the top quark often receives special attention in new physics models because its mass requires near unity coupling to the Higgs boson\cite{top}.
As shown in table \ref{SM table}, the top quark lifetime is so small that they decay before they can hadronise, that is to say, the process to create hadrons from quarks and gluons.\\

@@

The production of top quark can be generated in differents ways. In tevatron and LHC, the production of $t\bar{t}$ comes from ggF (gluon gluon fusion) via strong force. Even it is posible to create a single top quark in the LHC via electroweak interactions (with a W boson)\cite{th1}.  
The top quark, due to its big mass, always decays into a W boson and a b quark. But since the W boson have various decays, the top quark can generate many posible particles, acording to W boson decays. Some posible decays of top quark is the decay to a lepton such as muon, electron or tau and their neutrinos . %The antitop, antiparticle of top quark, can decay to W$^{-1} \bar{b}$ and its decay is the same as top, but with inverse signal.

@@

$W^i$ (i=1,2,3) are three  SU(2) gauge bosons and B is a U(1) gauge boson.\footnote{$\epsilon^{ijk}$ represents the Levi-Civita tensor, where $\epsilon^{ijk}= \begin{dcases}
	+1 \quad  \text{for even permutation} \\
	-1 \quad \text{for odd permutation}\\
	0 \quad \text{i=j, or j=k, or k=i } 
	\end{dcases}$.} In the end, those gauge bosons, after introduce the Higgs mechanism, become the $W^\pm$ , $Z$ and $\gamma$ vector boson
\cite{ew1}\cite{ew2}. %Z bosons  and photons will be formed by a combination of W and B fields.\\
\\

@@

In the recent years, people have studied the top anti top Higgs  ($t\bar{t}H$) process, in order to detect a Higgs boson, via multilepton signals\cite{th1}.
In this process, the creation of a Higgs boson comes from a intection via $t\bar{t}$. In the figure \ref{psu}, the $t\bar{t}H$ starts with 2 gluons which decay to two $t\bar{t}$ pairs and after that , the interaction $t\bar{t}$  generates a Higgs. The last process, subject of study in this thesis, is the top Higgs ($tH$) process. This process is a rare expected process with a very small production rate\cite{pd}.  In the figure \ref{thw}, is another way to produce $tH$, in association with a $W$.

@@

The number of particles decreases over time exponentially and this can be measured. With this, it is possible determinate the mean lifetime\cite{griff}
\begin{align}
\tau=\frac{1}{\Gamma}
\end{align}
In case of  the decay rate to a specific process, it is necessary get the branching fraction of the process. 
The branching ratio for a decay process is the ratio of the number of particles which decay via a specific decay mode with respect to the total number of particles which decay via all decay modes.
\begin{align}
BR_i =\frac{\Gamma_i}{\sum_{i}\Gamma_i}
\end{align}
Where $\Gamma=\sum_i\Gamma_i$ is the total decay width (sum of all partial widths) of the particle and is related to lifetime of the particle: $\Gamma=1/\tau$.
Since the dimension of $\Gamma$ is the inverse of time, in our system of natural units, it is measured in inverse seconds,it has the same dimension as mass (or energy)\footnote{Cleaves H.J. (2011) Branching Ratio. In: Gargaud M. et al. (eds) Encyclopedia of Astrobiology. Springer, Berlin, Heidelberg} . The lifetime of the Higgs boson is predicted to be 1.56 $\times$ $10^{-22}$ seconds and corresponds to $\Gamma_H$ is about 4 MeV , but has not yet been measured due to the detector resolution\cite{cms-manual}.
\\
\begin{table}[!htbp] 
\caption{SM Higgs boson branching ratios for  $M_H$ =125 GeV \protect \cite{pd}}
\centering
\begin{tabular}{|c|c|}
\hline
Higgs decay & Branching ratio (BR)\\
\hline
$H \rightarrow$ b$\bar{b}$ &$58.4\%$ \\
\hline
 $H \rightarrow$ $W^+W^-$ &$21.4\%$ \\
\hline
$H \rightarrow$ $\tau^+ \tau^-$ & $6.27\%$\\
\hline
$H \rightarrow$ ZZ &$2.62\%$\\
\hline
$H \rightarrow$ $\gamma\gamma$ &$0.227\%$\\
\hline
$H \rightarrow$ Z$\gamma$ &$0.153\%$\\
\hline
$H \rightarrow$ $\mu^+\mu^-$ &$0.0218\%$\\
\hline
\end{tabular}
\label{higgs1}
\end{table}
The decay  $H \rightarrow$ $b\bar{b}$ have a branching ratio a bit more than 50$\%$, therefore is the most likely decay Higgs boson can decay. Due to this,  $H\rightarrow b\bar{b}$ is a very studied process and most likely to detect a Higgs boson in the data after reconstruction. The decay  $H \rightarrow$ $\mu^+\mu^-$ has a very low branching ratio. This decay is very rare but not imposible to detect in the future.  In this work the majority of the Higgs decays detected are from $WW$, $ZZ$, $\tau \tau$\\

### 2019-10-15 tesis_hiram/Chapter2/chapter2.tex
algo

\begin{figure}[!htbp]
\centering
\includegraphics[scale=0.5]{Chapter2/lum6.png}
\caption{LHC schedule, including future plans for increase the center of mass energy and luminosity. The periods labeled LS are the maintenance periods \cite{cern3}.}
\label{lhc-lumi}
\end{figure}
From the figure \ref{lhc-lumi}, shows the Large Hadron Collider forecast for the increase of luminosity for the next years. Red dots are peak luminosity expected to reach and blue line is integrated luminosity. In the year 2018, the maximum integrated luminosity is around 150 $fb^{-1}$\cite{cern3}. From 2019 to 2021 the second phase of maintenance, where engineers increase the performance of the accelerator, give maintenance to the system and introduce new components to the complex. After that period, the LHC starts new collision period at even higher energies in order to explore new particle phenomena.

@@

%The produced muons are measured in three zones: in the inner tracker, after the coil, and in
%the return flux. Measurement the momentum of muons in the muon detector is determined by the muon bending angle at the exit of the 4T coil with the interaction point as the origin of the muon\cite{cms7}\cite{cms-manual}.
%The resolution of the muom detector is based on simulations taking account with the geometry and a test beam. 
The main purpose of the Muon System Detector is to identify muon tracks. 
The efficiency of the muon system is estimated by using single muon samples simulated with $p_t$=10, 50, 100, 500 and 1000 GeV. The results are shown in the figure \ref{mu-efi}.
\begin{figure}
	\centering
	\includegraphics[width=0.4\linewidth]{Chapter2/mu-efi}
	\caption{Muon reconstruction efficiency in terms of $\eta$ and $p_t$ \cite{cms-manual}}
	\label{mu-efi}
\end{figure}
The graph shows an efficiency of approximately 98$\%$ in the the fully instrumented regions. In the intersection regions, there are drops in efficiency  due to the separation of the chambers. % There is a reduction of the efficiency to 6$\%$, which corresponds to the $\eta$  position for the separation.

>>>>

\begin{figure}[!htbp]
\centering
\includegraphics[scale=0.5]{Chapter2/lum6.png}
\caption{LHC schedule, including future plans for increase the center of mass energy and luminosity. The periods labeled LS are the maintenance periods\cite{cern3}.}
\label{lhc-lumi}
\end{figure}
From the figure \ref{lhc-lumi}, shows the Large Hadron Collider forecast for the increase of luminosity for the next years. Red dots are peak luminosity expected to reach and blue line is integrated luminosity. In the year 2018, the maximum integrated luminosity is around 150 $fb^{-1}$\cite{cern3}. From 2019 to 2021 the second phase of maintenance, where engineers increase the performance of the accelerator, give maintenance to the system and introduce new components to the complex. After that period, the LHC starts new collision period at even higher energies in order to explore new particle phenomena.

@@

%The produced muons are measured in three zones: in the inner tracker, after the coil, and in
%the return flux. Measurement the momentum of muons in the muon detector is determined by the muon bending angle at the exit of the 4T coil with the interaction point as the origin of the muon\cite{cms7}\cite{cms-manual}.
%The resolution of the muom detector is based on simulations taking account with the geometry and a test beam. 
The main purpose of the Muon System Detector is to identify muon tracks. 
The efficiency of the muon system is estimated by using single muon samples simulated with $p_t$=10, 50, 100, 500 and 1000 GeV. The results are shown in the figure \ref{mu-efi}.
\begin{figure}
	\centering
	\includegraphics[width=0.4\linewidth]{Chapter2/mu-efi}
	\caption{Muon reconstruction efficiency in terms of $\eta$ and $p_t$\cite{cms-manual}}
	\label{mu-efi}
\end{figure}
The graph shows an efficiency of approximately 98$\%$ in the the fully instrumented regions. In the intersection regions, there are drops in efficiency  due to the separation of the chambers. % There is a reduction of the efficiency to 6$\%$, which corresponds to the $\eta$  position for the separation.
